Quantile Regression Analysis for Statin Effects on Body mass Index

### 2021-05-10 manuscript/manuscript2.Rmd
Update

  Body Mass Index (BMI) plays an important role in predicting heart disease risk[@katbody]. Approximately, 18 million deaths annually per year are caused by cardiovascular diseases(CVD), and similar to this number of nonfatal cardiovascular events [@hay2017gbd].
In 2011, annual costs for CVD and stroke were $\$320.1$ billion, which is more than cancer cost. This cost includes \$195.6  billion in direct costs (healthcare costs), and the cost of future productivity loss is  \$124.5 [@mozaffarian2015executive]. Important factors associated with CVD are high body mass index (BMI) and abnormal lipid ratio  [@yusuf2004effect], [@anderson1991]. Lowering low-density lipoprotein (LDL) by using statin or other drugs reduces the risk of cardiovascular diseases even with a population with no CVD  [@yusuf2016]. Ferrieres et al. [@ferrieres2018] studied the effect of BMI on the choice of lipid-lowering treatment. It is been found that statin intensity increases with the high level of BMI($\rho=0.13$).

@@

\begin{equation}
y_{i}=  
 \left\{
\begin{array}{ll}
 \beta_{01}+\beta_{11}x_i+\beta_{21}x_i^2+ \beta_{31}x_i^3+	\epsilon_i & \mbox{if } x_i< c \\
 \beta_{02}+\beta_{12}x_i+\beta_{22}x_i^2+ \beta_{23}x_i^3+	\epsilon_i& \mbox{if } x_i \geq c
\end{array}
\right.
\end{equation}
Imposing a continuty condition  and first, and second derivative at $c$ are equal in the two sides, we get a cubic splines. 
<!-- \begin{equation} -->
<!-- f(X)=X\beta=\beta_{0}+\beta_{1}X+\beta_2X^2+\beta_{3}X^{3}+\beta_{4}(X-a)_{+}^3+\beta_{5}(X-b)_{+}^3+\beta_6(X-c)_{+}^3, -->
<!-- \end{equation} -->

@@

 \begin{equation}
 \hat{BMI}= Total_chol+ Total_chol^2+ Race + Gender + Cholesterol_Drug_Use
 \end{equation}

\section{Numerical Example}
\subsection{Data}
 The data used in this study is National Health and Nutrition Examination Survey data (NHANES)[@NHANES]. The survey examines a nationally representative sample of the U.S. population. It focuses on a variety of health and nutrition measurements. The survey data are released every two years cycle.  In this study, we accumulated 6 cycles of NHANES data (2007–2018). We used two data files: One contains demographic variables, such as age, sex, race, income, etc., and the other contains data that are related to body measurements, such as BMI, head circumference, etc. These files are merged by using the respondent sequence number (SEQN)
 There are around 12,000 records. We selected a population age between 20 and 80.
 BMI are classified into different categories according to underweight, 18.5 kg/m2; normal weight, 18.5 to 25 kg/m2; overweight, 25 to 30 kg/m2; obese, 30 to 35 kg/m2; and very obese more than 35 kg/m2.

@@

@@

On the other hand, when we use splines to modeled the total cholesterol, there is a huge difference in the shape of the association expect at the higher quantile there is some similarity. At the lower quantile, there is negative correlation between total cholesterol and BMI for total cholesterol in the range less than 160, then the correlation becomes positive for total cholesterol in the range of (170, 250). at the 55th quantile, the correlation is positive for low cholesterol level and then becomes almost flat for the rest cholesterol values. At a higher qunatile, the correlation becomes negative at cholesterol level in the range (160,300), see Figure \@ref(resu330).

\section{Conclusion}
 Multivariate quantile regression is used to study the effects of different risk factors on the BMI levels. Cholesterol drug effects on BMI is negligible at low quantile but at higher quantile cholesterol medication  effects on BMI is larger. This study showed that the association between BMI and total cholesterol is varying with repect to different quantile.

>>>>

  Body Mass Index (BMI) plays an important role in predicting heart disease risk[@katbody]. Approximately, 18 million deaths annually per year are caused by cardiovascular diseases(CVD), and similar to this number of nonfatal cardiovascular events [@hay2017gbd].
In 2011, annual costs for CVD and stroke were $\$320.1$ billion, which is more than cancer cost. This cost includes \$195.6  billion in direct costs (health-care costs), and the cost of future productivity loss is  \$124.5 [@mozaffarian2015executive]. Important factors associated with CVD are high body mass index (BMI) and abnormal lipid ratio  [@yusuf2004effect], [@anderson1991]. Lowering low-density lipoprotein (LDL) by using statin or other drugs reduces the risk of cardiovascular diseases even with a population with no CVD  [@yusuf2016]. Ferrieres et al. [@ferrieres2018] studied the effect of BMI on the choice of lipid-lowering treatment. It is been found that statin intensity increases with the high level of BMI($\rho=0.13$).

@@

\begin{equation}
y_{i}=  
 \left\{
\begin{array}{ll}
 \beta_{01}+\beta_{11}x_i+\beta_{21}x_i^2+ \beta_{31}x_i^3+	\epsilon_i & \mbox{if } x_i< c \\
 \beta_{02}+\beta_{12}x_i+\beta_{22}x_i^2+ \beta_{23}x_i^3+	\epsilon_i& \mbox{if } x_i \geq c
\end{array}
\right.
\end{equation}
Imposing a continuity condition  and first, and second derivative at $c$ are equal in the two sides, we get a cubic splines. 
<!-- \begin{equation} -->
<!-- f(X)=X\beta=\beta_{0}+\beta_{1}X+\beta_2X^2+\beta_{3}X^{3}+\beta_{4}(X-a)_{+}^3+\beta_{5}(X-b)_{+}^3+\beta_6(X-c)_{+}^3, -->
<!-- \end{equation} -->

@@

 \begin{equation}
 \hat{BMI}= Total chol+ Total chol^2+ Race + Gender + Cholesterol Drug_Use
 \end{equation}

\section{Numerical Example}
\subsection{Data}
 The data used in this study is the National Health and Nutrition Examination Survey data (NHANES)[@NHANES]. The survey examines a nationally representative sample of the U.S. population. It focuses on a variety of health and nutrition measurements. The survey data are released every two years cycle.  In this study, we accumulated 6 cycles of NHANES data (2007–2018). We used two data files: One contains demographic variables, such as age, sex, race, income, etc., and the other contains data that are related to body measurements, such as BMI, head circumference, etc. These files are merged by using the respondent sequence number (SEQN)
 There are around 12,000 records. We selected a population age between 20 and 80.
 BMI are classified into different categories according to underweight, 18.5 kg/m2; normal weight, 18.5 to 25 kg/m2; overweight, 25 to 30 kg/m2; obese, 30 to 35 kg/m2; and very obese more than 35 kg/m2.

@@

```{r resu1000, out.width = "100%", fig.align = "center", fig.cap = "  An illustration of  marginal effects of age on different BMI quantiles. The predictors modeled using quadratic terms and splines.", echo = FALSE}
# Created by file exploratory-data-analysis.R
if (!file.exists(here::here("images", "comp_age_mixed")))  
knitr::include_graphics(here::here("images", "comp_age_mixed.png"))
```

```{r resu1010, out.width = "100%", fig.align = "center", fig.cap = " An illustration of  marginal effects of total cholesterol  on different BMI quantiles. The predictors modeled using quadratic terms and splines.", echo = FALSE}
# Created by file exploratory-data-analysis.R
if (!file.exists(here::here("images", "comp_tot_mixed")))  
knitr::include_graphics(here::here("images", "comp_tot_mixed.png"))
```

@@

On the other hand, when we use splines to modeled the total cholesterol, there is a huge difference in the shape of the association expect at the higher quantile there is some similarity. At the lower quantile, there is negative correlation between total cholesterol and BMI for total cholesterol in the range less than 160, then the correlation becomes positive for total cholesterol in the range of (170, 250). at the 55th quantile, the correlation is positive for low cholesterol level and then becomes almost flat for the rest cholesterol values. At a higher quntile, the correlation becomes negative at cholesterol level in the range (160,300), see Figure \@ref(resu330).

\section{Conclusion}
 Multivariate quantile regression is used to study the effects of different risk factors on the BMI levels. Cholesterol drug effects on BMI is negligible at low quantile but at higher quantile cholesterol medication  effects on BMI is larger. This study showed that the association between BMI and total cholesterol is varying with respect to different quantile.

### 2021-05-10 manuscript/manuscript2.tex
Update

\title{Quantile Regression Analysis for Statin Effects on Body mass Index}
\author{}
\date{\vspace{-2.5em}April 22, 2021}

@@

Body Mass Index (BMI) plays an important role in predicting heart disease risk(Katzmarzyk et al. 2012). Approximately, 18 million deaths annually per year are caused by cardiovascular diseases(CVD), and similar to this number of nonfatal cardiovascular events (Hay et al. 2017).
In 2011, annual costs for CVD and stroke were \(\$320.1\) billion, which is more than cancer cost. This cost includes \$195.6 billion in direct costs (healthcare costs), and the cost of future productivity loss is \$124.5 (Mozaffarian et al. 2015). Important factors associated with CVD are high body mass index (BMI) and abnormal lipid ratio (Yusuf et al. 2004), (Anderson et al. 1991). Lowering low-density lipoprotein (LDL) by using statin or other drugs reduces the risk of cardiovascular diseases even with a population with no CVD (Yusuf et al. 2016). Ferrieres et al.~(Ferrières et al. 2018) studied the effect of BMI on the choice of lipid-lowering treatment. It is been found that statin intensity increases with the high level of BMI(\(\rho=0.13\)).

@@

\begin{equation}
y_{i}=  
 \left\{
\begin{array}{ll}
 \beta_{01}+\beta_{11}x_i+\beta_{21}x_i^2+ \beta_{31}x_i^3+ \epsilon_i & \mbox{if } x_i< c \\
 \beta_{02}+\beta_{12}x_i+\beta_{22}x_i^2+ \beta_{23}x_i^3+ \epsilon_i& \mbox{if } x_i \geq c
\end{array}
\right.
\end{equation}
Imposing a continuty condition and first, and second derivative at \(c\) are equal in the two sides, we get a cubic splines.

@@

\begin{equation}
 \hat{BMI}= Total_chol+ Total_chol^2+ Race + Gender + Cholesterol_Drug_Use
 \end{equation}

@@

The data used in this study is National Health and Nutrition Examination Survey data (NHANES)(Disease Control and (CDC) 2018). The survey examines a nationally representative sample of the U.S. population. It focuses on a variety of health and nutrition measurements. The survey data are released every two years cycle. In this study, we accumulated 6 cycles of NHANES data (2007--2018). We used two data files: One contains demographic variables, such as age, sex, race, income, etc., and the other contains data that are related to body measurements, such as BMI, head circumference, etc. These files are merged by using the respondent sequence number (SEQN)
There are around 12,000 records. We selected a population age between 20 and 80.
BMI are classified into different categories according to underweight, 18.5 kg/m2; normal weight, 18.5 to 25 kg/m2; overweight, 25 to 30 kg/m2; obese, 30 to 35 kg/m2; and very obese more than 35 kg/m2.

@@

\caption{Quantile regression of BMI ploted f for fifferent marginal effects. The predictors modeled using quadratic terms }\label{fig:resu1}
\end{figure}

@@

\caption{Quantile regression of BMI ploted f for fifferent marginal effects. The predictors modeled using splines }\label{fig:resu12}
\end{figure}

@@

@@

On the other hand, when we use splines to modeled the total cholesterol, there is a huge difference in the shape of the association expect at the higher quantile there is some similarity. At the lower quantile, there is negative correlation between total cholesterol and BMI for total cholesterol in the range less than 160, then the correlation becomes positive for total cholesterol in the range of (170, 250). at the 55th quantile, the correlation is positive for low cholesterol level and then becomes almost flat for the rest cholesterol values. At a higher qunatile, the correlation becomes negative at cholesterol level in the range (160,300), see Figure \ref{resu330}.

@@

Multivariate quantile regression is used to study the effects of different risk factors on the BMI levels. Cholesterol drug effects on BMI is negligible at low quantile but at higher quantile cholesterol medication effects on BMI is larger. This study showed that the association between BMI and total cholesterol is varying with repect to different quantile.

>>>>

\title{Quantile Regression Analysis for Statin Effects on Body mass Index}
\author{}
\date{\vspace{-2.5em}May 10, 2021}

@@

Body Mass Index (BMI) plays an important role in predicting heart disease risk(Katzmarzyk et al. 2012). Approximately, 18 million deaths annually per year are caused by cardiovascular diseases(CVD), and similar to this number of nonfatal cardiovascular events (Hay et al. 2017).
In 2011, annual costs for CVD and stroke were \(\$320.1\) billion, which is more than cancer cost. This cost includes \$195.6 billion in direct costs (health-care costs), and the cost of future productivity loss is \$124.5 (Mozaffarian et al. 2015). Important factors associated with CVD are high body mass index (BMI) and abnormal lipid ratio (Yusuf et al. 2004), (Anderson et al. 1991). Lowering low-density lipoprotein (LDL) by using statin or other drugs reduces the risk of cardiovascular diseases even with a population with no CVD (Yusuf et al. 2016). Ferrieres et al.~(Ferrières et al. 2018) studied the effect of BMI on the choice of lipid-lowering treatment. It is been found that statin intensity increases with the high level of BMI(\(\rho=0.13\)).

@@

\begin{equation}
y_{i}=  
 \left\{
\begin{array}{ll}
 \beta_{01}+\beta_{11}x_i+\beta_{21}x_i^2+ \beta_{31}x_i^3+ \epsilon_i & \mbox{if } x_i< c \\
 \beta_{02}+\beta_{12}x_i+\beta_{22}x_i^2+ \beta_{23}x_i^3+ \epsilon_i& \mbox{if } x_i \geq c
\end{array}
\right.
\end{equation}
Imposing a continuity condition and first, and second derivative at \(c\) are equal in the two sides, we get a cubic splines.

@@

\begin{equation}
 \hat{BMI}= Total chol+ Total chol^2+ Race + Gender + Cholesterol Drug_Use
 \end{equation}

@@

The data used in this study is the National Health and Nutrition Examination Survey data (NHANES)(Disease Control and (CDC) 2018). The survey examines a nationally representative sample of the U.S. population. It focuses on a variety of health and nutrition measurements. The survey data are released every two years cycle. In this study, we accumulated 6 cycles of NHANES data (2007--2018). We used two data files: One contains demographic variables, such as age, sex, race, income, etc., and the other contains data that are related to body measurements, such as BMI, head circumference, etc. These files are merged by using the respondent sequence number (SEQN)
There are around 12,000 records. We selected a population age between 20 and 80.
BMI are classified into different categories according to underweight, 18.5 kg/m2; normal weight, 18.5 to 25 kg/m2; overweight, 25 to 30 kg/m2; obese, 30 to 35 kg/m2; and very obese more than 35 kg/m2.

@@

\caption{Quantile regression of BMI ploted  for different marginal effects. The predictors modeled using quadratic terms }\label{fig:resu1}
\end{figure}

@@

\caption{Quantile regression of BMI ploted  for different marginal effects. The predictors modeled using splines }\label{fig:resu12}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{/home/mabdulsa/BMI/images/comp_age_mixed}

\caption{  An illustration of  marginal effects of age on different BMI quantiles. The predictors modeled using quadratic terms and splines.}\label{fig:resu1000}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{/home/mabdulsa/BMI/images/comp_tot_mixed}

\caption{ An illustration of  marginal effects of total cholesterol  on different BMI quantiles. The predictors modeled using quadratic terms and splines.}\label{fig:resu1010}
\end{figure}

@@

On the other hand, when we use splines to modeled the total cholesterol, there is a huge difference in the shape of the association expect at the higher quantile there is some similarity. At the lower quantile, there is negative correlation between total cholesterol and BMI for total cholesterol in the range less than 160, then the correlation becomes positive for total cholesterol in the range of (170, 250). at the 55th quantile, the correlation is positive for low cholesterol level and then becomes almost flat for the rest cholesterol values. At a higher quntile, the correlation becomes negative at cholesterol level in the range (160,300), see Figure \ref{resu330}.

@@

Multivariate quantile regression is used to study the effects of different risk factors on the BMI levels. Cholesterol drug effects on BMI is negligible at low quantile but at higher quantile cholesterol medication effects on BMI is larger. This study showed that the association between BMI and total cholesterol is varying with respect to different quantile.
